---
title: "High Lonesome 2021 Lottery"
author: "Garret Christensen"
date: "January 13, 2021"
output: html_document
runtime: shiny
---
## Background

Welcome to the High Lonesome 2021 Lottery. This lottery is run with code written by Garret Christensen. The code is available on GitHub [here](http://www.github.com/garretchristensen/RaceLottery2021). The intent with using Github is to make the entire process transparent and reproducible. If you speak R, please check it out!

## Lottery Design Choices
The 2021 lottery is, like everything else, drastically affected by the pandemic and the cancellation of the 2020 race. We are mostly filling the slots of runners who chose not to roll over their 2020 entry. We are picking 5 men, 14 women, and a 15 person waitlist, starting with women and then alternating gender. 

The general goals of the lottery are as follows:

* We want equal numbers of men and women
* We'd like to get a mix of new and veteran runners, without guaranteeing entry for either
* Previous unsuccessful applications should be the major determinant of selection
* We value volunteering and trail work
* We'd like new entrants to have a decent chance to run within a couple-few years

So here are the activities for which we will award points:

* Volunteer shifts *at* High Lonesome
* Extra volunteer trailwork beyond the eight hours required
* Previous applications for the race
* Previous finishes of the race


### Chosen Model

Statistically, getting equal numbers of men and women is normally easy---just divide the race cap in half and draw separately. The number of slots drawn in 2021 is different because of the number of 2020 rollovers for each gender. The rest is a bit more complicated, but we've chosen the following weighting method:

Tickets=2^(n+k+1)+2ln(v+t+1) where n, k, v, and t are defined as follows:

* Previous unsuccessful applications: Since you were last *picked* in the lottery, offered entry off the waitlist, or offered entry via the volunteer raffle, how many times have you entered the High Lonesome lottery, not including this year? Note that previous lottery entries need not be in successive years, but they reset to zero after you are picked in the lottery or are offered a slot via other means such as the waitlist or the volunteer raffle. n is equal to your previous unsuccessful applications, as defined above, with no maximum. For 2021 this is at most 1 because we've only had one lottery before.


* Previous finishes: How many times have you finished High Lonesome 100?
We will award previous finishers a boost until they have finished High Lonesome three times; then the boost decreases. (We love returning runners, but after a while, it's cool to give others a chance.)
k is defined according to the following rule:
* k=0 if finishes==0
* k=0.5 if finishes==1
* k=1 if finishes==2
* k=1.5 if finishes==3
* k=0.5 if finishes>=4

* Volunteer shifts: How many official shifts did you volunteer at the previous (current calendar year at the time of the lottery) running of High Lonesome? v is  the number of official shifts. Maximum of 10. 

* Trailwork: Not including the trailwork *required* for High Lonesome, how many solid 8-hour shifts of physical volunteer trail work have you completed since November 15 of last year? Volunteering or pacing at a race doesn't count. These trailwork hours must be *over and above* the 8-hours required for High Lonesome itself, they must be done with an approved land management agency or partner organization, and they must not be used in order to qualify for any other race or ultra--or court-mandated community service requirement ;). We obviously don't have a great way to verify this, so you're on your honor. Please don't abuse this. t is the number of 8-hour shifts. Maximum of 10.

Note that runners are allowed deferred entry from previous years due to pregnancy or new children or getting into Western States or Hardrock. There are also a small number of legacy spots. We take these slots from their respective gender lotteries. All these policies are described on the High Lonesome webpage [here](https://www.highlonesome100.com/general-registration-info).


```{r prep, echo=FALSE, results=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(tibble)
library(readxl)


#####################################################
#SET EVERYTHING UP BEFORE DOING THE SHINY PART
#######################################################

#I COULD MAKE THIS AN INTERACTIVE PART TOO LATER

temp<-read.csv("./2021HL100FINAL.csv", stringsAsFactors = FALSE) #LOAD THE DATA
df<-as_tibble(temp)
#THERE IS SOME STUPID NON-ASCII CHARACTER BEGINNING-OF-FILE 
#NONSENSE GOING ON HERE
#CHEAP WORKAROUND IS JUST TO MAKE SURE FIRST_NAME IS NOT THE FIRST VARIABLE!
#names(df) <- iconv(names(df), to='ASCII', sub='')
#df$First_Name<-df$..First_Name

df$fullname<-paste(df$First_Name, df$Last_Name, sep=" ", collapse = NULL)
head(df)

#df$USA<-ifelse(df$Country=="USA",1,0)

#NUMBER OF MEN AND WOMEN APPLICANTS
n_men_app=nrow(men<-df[which(df$Gender=="M"),])
n_women_app=nrow(women<-df[which(df$Gender=="F"),])

```

# Implementing the Lottery
To run the lottery, all you have to do is enter the seed below. (It will show an error until you enter a number.) The actual seed will be determined by a live roll of nine 10-sided dice on Friday January 15, each single die roll becoming a digit of the integer, in order, plus one coin toss determining whether the entire number is positive or negative. The seed primes R's pseudo-random number generator so the code can run reproducibly. Before the seed is set, applicants can enter any integer they want to simulate the lottery. Well, almost. For computer-y reasons it can be any integer with absolute value less than or equal to 2,147,483,647.



```{r runlottery, echo=FALSE, warning=FALSE}
shinyApp(
  
#######################################################################
# WHAT THE UI LOOKS LIKE
#######################################################################
ui<-fluidPage(
  # Copy the line below to make a number input box into the UI.
  fluidRow(
  column(6, numericInput("num", label = h1("Enter the seed"), value = NA))
  ),
  
  #fluidRow(
  #column(6, numericInput("w_defer", label = h1("Number of female deferrals/legacy runners"), value = 0)),
  #column(6, numericInput("m_defer", label = h1("Number of male deferrals/legacy runners"), value = 5)) 
  #),             
  
  fluidRow(verbatimTextOutput("value")),
  #what do I have to do to get this printed above the data tables?
  fluidRow("These are the women selected in the lottery:", tableOutput("women")),
  #and something inbetween them would be nice.
  fluidRow("These are the men selected in the lottery:", tableOutput("men")),
  
  fluidRow("This is the waitlist:", tableOutput("combo"))
 ),

#############################################################################
server<- function(input, output) {

output$value <- renderPrint({
set.seed(input$num) #SET THE SEED WITH DICE!
nm<-input$Nm #SET THE TOTAL NUMBER TO PICK
nw<-input$Nw #SET THE TOTAL NUMBER TO PICK

######################################################################
#FOR 2021 JUST SET HOW MANY TO PICK
n_women_pick <- 14
n_men_pick <- 5
      
######################################################################
#DETERMINE TICKETS FROM THE DATA
#For 2020 nobody has unsuccessful applications, this is the first lottery
df$Applications<-df$Previous_Applications

#k is defined according to the following rule:
# k=0 if finishes==0
#k=0.5 if finishes==1
#k=1 if finishes==2
#k=1.5 if finishes==3
#k=1 if finishes>=4
df$k <- ifelse(df$Previous_Finishes==0 , 0,
               ifelse(df$Previous_Finishes==1,  0.5,
                      ifelse(df$Previous_Finishes==2, 1, 
                             ifelse(df$Previous_Finishes==3, 1.5,
                                    ifelse(df$Previous_Finishes>=4, 0.5, 0)))))


#Shifts max out at 10
df$n<-pmin(df$Volunteer_Shifts, 10)
df$t<-pmin(df$Trailwork, 10)

#Tickets=2^(n+k+1)+2ln(v+t+1) where n, k, v, and t are defined as follows:
df$tickets <-2^(df$k+df$Applications+1) + 2*log(df$n+df$t+1)

#SPLIT THE DATA INTO MENS AND WOMENS
men<-df[which(df$Gender=="M"),]
women<-df[which(df$Gender=="F"),]

##############################################################
#DRAW THE LOTTERY

    #dplyr function sample_n will work with weights, normalize automatically
    #syntax:sample_n(tbl, size, replace = FALSE, weight = NULL, .env = NULL, ...)
    #Run the separate lotteries
    women_winners<-sample_n(women, n_women_pick, replace = FALSE, weight=women$tickets)
    men_winners<-sample_n(men, n_men_pick, replace = FALSE, weight=men$tickets)
    
    #I can't figure out how to label tables, so just make the table itself
    #look sort of well-labeled
    #subset
    #dfnew5 <- subset(diamonds, select=c("color", "carat", "price"))
    w_output<-subset(women_winners, select=c("fullname"))
    m_output<-subset(men_winners, select=c("fullname"))
    
    #Add ID column data$ID <- seq.int(nrow(data))
    w_output$Num<-seq.int(nrow(w_output))
    m_output$Num<-seq.int(nrow(m_output))
    #rearrange columns df2[,c(1,3,2,4)]
    #w_output<-w_output[,c(5,1,2,3,4)]
    #m_output<-m_output[,c(5,1,2,3,4)]
    #rename: names(data)[3]<-"new_name"
    names(w_output)[1]<-"Selected_Women"
    names(m_output)[1]<-"Selected_Men"
    #Send the winners' names to be output
    output$women <- renderTable(w_output)
    output$men <- renderTable(m_output)
      
      
  #########################################################
  #waitlist
  #########################################################

    women_waitlist_pool<-anti_join(women, women_winners)
    n_women_waitlist_pool<-nrow(women_waitlist_pool)
 
  men_waitlist_pool<-anti_join(men, men_winners)
  n_men_waitlist_pool<-nrow(men_waitlist_pool)
    
  #SIMPLER THIS YEAR, JUST ENTER THE NUMBERS FOR THE WL, 8 and 7
  n_women_wait_pick<-13
  n_men_wait_pick<-12
  
  #PICK THE WAITLISTERS
  #WOMEN MIGHT NOT HAVE ANY

    women_waiters <- sample_n(women_waitlist_pool, n_women_wait_pick, replace = FALSE, weight=women_waitlist_pool$tickets)
    w_output_wait<-subset(women_waiters, select=c("fullname"))
    w_output_wait_priv<-subset(women_waiters, select=c("fullname", "Email_Address"))
    w_output_wait$Num<-seq.int(nrow(w_output_wait))
    #w_output_wait<-w_output_wait[,c(5,1,2,3,4)]
    names(w_output_wait)[1]<-"Waitlisted_Women"
    output$women_wait <- renderTable(w_output_wait)
  
  #ASSUME MEN WILL HAVE ENOUGH FOR A FULL WAITLIST
  men_waiters <- sample_n(men_waitlist_pool, n_men_wait_pick, replace = FALSE, weight=men_waitlist_pool$tickets)
      
#I can't figure out how to label tables, so just make the table itself
  #look sort of well-labeled
  #subset
  #dfnew5 <- subset(diamonds, select=c("color", "carat", "price"))
  m_output_wait<-subset(men_waiters, select=c("fullname"))
  m_output_wait_priv<-subset(men_waiters, select=c("fullname", "Email_Address"))
  #Add ID column data$ID <- seq.int(nrow(data))
  m_output_wait$Num<-seq.int(nrow(m_output_wait))
  #rearrange columns df2[,c(1,3,2,4)
  #m_output_wait<-m_output_wait[,c(5,1,2,3,4)]
  #rename: names(data)[3]<-"new_name"
  names(m_output_wait)[1]<-"Waitlisted_Men"
  #Send the winners' names to be output
    output$men_wait <- renderTable(m_output_wait)   
  
      
  ##########################################
  #Zipper the waitlists
  ########################################
  #make column names identical so columns line up
  names(m_output_wait)[1]<-"Waitlisted_Name"
  names(w_output_wait)[1]<-"Waitlisted_Name"
  #bind women first for the waitlist for 2021
  
  temp <- bind_rows(w_output_wait, m_output_wait)
  
  #sort and relabel
  temp <- arrange(temp, Num)
  temp$GenderNum <-temp$Num
  temp$Num <-seq.int(nrow(temp))
  
  output$combo<-renderTable(temp)
   
})
},
 options = list(height = 900)
)

```


### 2021 Waitlist
We picked 25 people for the waitlist, starting with women.
